---
title: "Bike_SharingTripHistory_Analysis"
author: "Swagata Kundu"
date: "7/16/2017"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(readr)
Q1_triphistory_2011 <- read_csv("~/swag_projects/Capstone_project/BikeRental_Deman_Prediction/2011-Q1-cabi-trip-history-data.csv", 
    col_types = cols(Enddate = col_datetime(format = "%m/%d/%Y %H:%M"), 
        Startdate = col_datetime(format = "%m/%d/%Y %H:%M")))


Q2_triphistory_2011 <- read_csv("~/swag_projects/Capstone_project/BikeRental_Deman_Prediction/2011-Q2-cabi-trip-history-data.csv", 
    col_types = cols(Enddate = col_datetime(format = "%m/%d/%Y %H:%M"), 
        Startdate = col_datetime(format = "%m/%d/%Y %H:%M")))


Q3_triphistory_2011 <- read_csv("~/swag_projects/Capstone_project/BikeRental_Deman_Prediction/2011-Q3-cabi-trip-history-data.csv", 
    col_types = cols(`Enddate` = col_datetime(format = "%m/%d/%Y %H:%M"), 
        `Startdate` = col_datetime(format = "%m/%d/%Y %H:%M")))



Q4_triphistory_2011 <- read_csv("~/swag_projects/Capstone_project/BikeRental_Deman_Prediction/2011-Q4-cabi-trip-history-data.csv", 
    col_types = cols(`Enddate` = col_datetime(format = "%m/%d/%Y %H:%M"), 
        `Startdate` = col_datetime(format = "%m/%d/%Y %H:%M")))


```

creating a new variable "data" for Q1_triphistory_2011 dataset and installing qdapRegex
for station numbers(start & end) extraction from entire station address
```{r}
data = rbind(Q1_triphistory_2011, Q2_triphistory_2011, Q3_triphistory_2011, Q4_triphistory_2011)

         
#install.packages("qdapRegex")
library(qdapRegex)

data$St_Stationnumber = gsub("[\\(\\)]", "", regmatches(data$Startstation, gregexpr("\\(.*?\\)", data$Startstation)))

data$End_Stationnumber = gsub("[\\(\\)]", "", regmatches(data$Endstation, gregexpr("\\(.*?\\)", data$Endstation)))
```

Uploading master station terminals with lattitude longitude coordinates to get the street address and zipcode(start & end) correspondent to station numbers and saving in variable "zipcodes"
To extract street address and zipcodes from latitude and longitude coordinates, we use ggplot2 & ggmap packages
```{r}

StationTerminalnos_latlong <- read_csv("~/swag_projects/Capstone_project/BikeRental_Deman_Prediction/Capital_Bike_Share_Locations.csv")


zipcodes = StationTerminalnos_latlong

#install.packages("ggplot2")
library(ggplot2)
#install.packages("ggmap")
library(ggmap)

zipcodes = zipcodes[,-c(7:16)]
  
location = do.call(rbind,
                   lapply(1:nrow(zipcodes),
                          function(i)revgeocode(as.numeric(zipcodes[i,6:5]))))



zipcodes = cbind(zipcodes,location)

#install.packages("stringr")
library(stringr)
zipcodes$zipcode <- substr(str_extract(zipcodes$location," [0-9]{5}, .+"),2,6)

write.csv(zipcodes, file = "new_bikelocation.csv")
```


Using SQLDF to join multiple dataset sttaion terminal with corresponding zipcodes, In this joining process less than 1% data is lost which might be due to NA's
```{r}

new_bikelocation <- read_csv("~/swag_projects/Capstone_project/BikeRental_Deman_Prediction/new_bikelocation.csv")


Publicholidays <- read_csv("~/swag_projects/Capstone_project/BikeRental_Deman_Prediction/Publicholidays_2011_2012.csv", 
    col_types = cols(Date = col_date(format = "%m/%d/%Y")))



data$sdate = as.Date(data$Startdate)


Bike_trip_hist = sqldf("Select data.MemberType, data.sdate, data.Startdate,  data.St_Stationnumber ,
                               new_bikelocation.zipcode as 'st_zipcode', data.Enddate, data.End_Stationnumber, x.zipcode as 'end_zipcode' , data.Duration , data.Bike_ID, 
CASE WHEN date(data.sdate == holidays.Date)
      THEN 1
      ELSE 0
END AS Holiday
                        From data
                        inner join new_bikelocation on data.St_Stationnumber =  new_bikelocation.TERMINAL_NUMBER
                      INNER JOIN new_bikelocation as x on data.End_Stationnumber = x.TERMINAL_NUMBER
                      LEFT OUTER JOIN Publicholidays as holidays on date(data.sdate) = date(holidays.Date)")
```



To validate our join , we run the following code
```{r}
sqldf("Select distinct(Bike_trip_hist.sdate)
      from Bike_trip_hist
      where Bike_trip_hist.Holiday = 1")

write.csv(Bike_trip_hist, file = "Bike_triphistory_data.csv")
```




